---
title: "Bike Share Demand"
author: "Kinga"
date: "Thursday, February 26, 2015"
output: html_document
---
##Reading in the data

```{r}
df <- read.csv("train.csv")
```


##Examininig the data

```{r}
head(df)
str(df$datetime)
tail(df)
nrow(df)
names(df)
```

##Preparing the data

####Locating the missing rows


```{r}
#Finding the missing rows:
library(lubridate)
class(df$datetime)
df$datetime <- ymd_hms(df$datetime)
class(df$datetime)
head(df$datetime)
problemRows <- c()
#gap <- list()
for (i in 2:nrow(df)){
        if ((as.numeric(df$datetime[i]-df$datetime[i-1], units="hours")>1) & 
                    (as.numeric(df$datetime[i]-df$datetime[i-1], units="hours")< 14)) {
                problemRows <- c(problemRows,i)
                #gap[[i]]<-(df$datetime[i]-df$datetime[i-1])
        }
}
problemRows
#finding the gap at the problem rows
gap <- list()

for (i in problemRows){
        #print(i)
        gap[[as.character(i)]]<-(as.numeric(df$datetime[i]-df$datetime[i-1], units="hours"))
        
}
gap[[1]]
gap[[2]]
head(gap, length(problemRows))
#the class of the difference is

class(df$datetime[31]-df$datetime[30])
```
So, what is the lowest count, i.e. the lowest number of bikes rented?

```{r}
min(df$count)
```


At what time did the missing data points occur?

```{r}
for (i in problemRows){
        print(i)
        print(gap[[as.character(i)]])
        print(df[i-1,1])
        print(df[i, 1])
        
}

```

And the answer is that the missing data points always occur after midnight and before 6 am, with the exception of thte interval between 

```{r}
df$datetime[396]
```

and 

```{r}
df$datetime[397]


```

and the interval 

```{r}
df[3723,]
df[3724,]

```

but where the time difference is

```{r}
as.numeric(df$datetime[3724]-df$datetime[3723], units="hours")

```


So it seems reasonable to me that the missing data values correspond to the times when 
no one rented a bike and "count"=0.

So, I will insert the missing rows, by assuming that count=0 for all those times.
I will do this even for the stretch of 13 hours of missing data on 2011-01-17 and 
2011-01-18 **for now**.  

####Imputing values for the rows that are missing


```{r}
for (i in problemRows){
        new_row<-df[(i-1),]
        new_row$temp <- as.numeric(mean(df$temp[(i-1):i]))
        new_row$atemp <- as.numeric(mean(df$atemp[(i-1):i]))
        new_row$windspeed <- as.numeric(mean(df$windspeed[(i-1):i]))
        new_row$humidity <- as.integer(mean(df$humidity[(i-1):i]))
        new_row$casual <- 0
        new_row$registered <- 0
        new_row$count <- 0
        for (j in 1 : (gap[[as.character(i)]]-1)){
              new_row$datetime <- new_row$datetime +hms("1 0 0") 
              df <- rbind(df, new_row)
              
        }
        
        
}
library(dplyr)
complete_df <- arrange(df, datetime)

```





And checking for gaps of less than 14 hours, but more than 1 hour:

```{r}
new_problem_rows<-c()
for (i in 2:nrow(complete_df)){
        if ((as.numeric(complete_df$datetime[i]-complete_df$datetime[i-1], units="hours")>1) & 
                    (as.numeric(complete_df$datetime[i]-complete_df$datetime[i-1], units="hours")< 14)) {
                new_problem_rows <- c(new_problem_rows,i)
                #gap[[i]]<-(df$datetime[i]-df$datetime[i-1])
        }
}
new_problem_rows




```

```{r}
Jan2011 <- filter(complete_df, datetime < ymd_hms("2011-02-01 00:00:00"))
Jan2011_incomplete <- filter(df, datetime < ymd_hms("2011-02-01 00:00:00"))
Feb2011<-filter(complete_df, datetime < ymd_hms("2011-03-01 00:00:00") & datetime >=ymd_hms("2011-02-01 00:00:00"))
```


